---
title: "Automobile Customer Segmentation"
output: html_notebook
---

Import data set and read data set into data frame
```{r}
customers <- read.csv("C:/Users/Jacque de l'est/Documents/Datasets for Data Science/Automobile Company Customers.csv", stringsAsFactors = FALSE)
```

show data frame
```{r}
head(customers)
```

Show data frame shape and columns
```{r}
str(customers)
```

show descriptive statistics of data set
```{r}
summary(customers)
```

See missing values of data frame
```{r}
colSums(is.na(customers))
```

Fill missing values of Work_Experience with 0 and Family_Size with 1
```{r}
customers$Work_Experience[is.na(customers$Work_Experience)] <- 0
customers$Family_Size[is.na(customers$Family_Size)] <- 1
colSums(is.na(customers))
```

Check unique sum and values of data frame columns
```{r}
sapply(customers, function(x) length(unique(x)))
```
```{r}
unique(customers$Gender)
unique(customers$Ever_Married)
unique(customers$Graduated)
unique(customers$Profession)
unique(customers$Spending_Score)
unique(customers$Var_1)
```

Show sum of empty values
```{r}
sum(customers$Ever_Married == "")
sum(customers$Graduated == "")
sum(customers$Profession == "")
sum(customers$Var_1 == "")
```

Change empty values to NA values
```{r}
customers[customers == ""] <- NA
head(customers)
```

Fill NA values with proper values
```{r}
customers$Ever_Married[is.na(customers$Ever_Married)] <- "No"
customers$Graduated[is.na(customers$Graduated)] <- "No"
customers$Profession[is.na(customers$Profession)] <- "others"
customers$Var_1[is.na(customers$Var_1)] <- "Cat_0"
sum(is.na(customers))
```

Show distribution of categorical data
```{r}
library(ggplot2)
ggplot(customers, aes(x = Gender, fill = Gender)) + geom_bar()
ggplot(customers, aes(x = Ever_Married, fill = Ever_Married)) + geom_bar()
ggplot(customers, aes(x = Graduated, fill = Graduated)) + geom_bar()
ggplot(customers, aes(x = Profession, fill = Profession)) + geom_bar()
ggplot(customers, aes(x = Spending_Score, fill = Spending_Score)) + geom_bar()
ggplot(customers, aes(x = Var_1, fill = Var_1)) + geom_bar()
```

Show distribution of numerical columns
```{r}
# Distribution of Age
#histogram & mean line
ggplot(customers, aes(x=Age, fill=Age)) +
    geom_histogram(color = "black", fill = "white") + geom_vline(aes(xintercept=mean(Age)),
               color="red", linetype="dashed", size=1)


#density curve
ggplot(customers, aes(x=Age)) + geom_density()

#histogram and mean line overlaid with kernel density curve
ggplot(customers, aes(x=Age)) + 
    geom_histogram(aes(y=..density..),
                   colour="black", fill="white") +
    geom_density(alpha=0.2, fill="#FF6666") + geom_vline(aes(xintercept=mean(Age)),
               color="red", linetype="dashed", size=1)
```
```{r}
# Distribution of Work Experience
#histogram & mean line
ggplot(customers, aes(x=Work_Experience, fill=Work_Experience)) +
    geom_histogram(binwidth = 1, color = "black", fill = "white") + geom_vline(aes(xintercept=mean(Work_Experience)),
               color="red", linetype="dashed", size=1)
```
```{r}
# Distribution of Family Size
#histogram & mean line
ggplot(customers, aes(x=Family_Size, fill=Family_Size)) +
    geom_histogram(binwidth = 1, color = "black", fill = "white") + geom_vline(aes(xintercept=mean(Family_Size)),
               color="red", linetype="dashed", size=1)
```

Drop column ID
```{r}
customers_new <- customers[-1]
head(customers_new)
```

Convert categorical columns to numerical columns
```{r}
cat_columns <- data.matrix(customers_new[c("Gender","Ever_Married","Graduated","Profession","Spending_Score","Var_1")])
head(cat_columns)
```

Merge converted categorical columns with original data frame
```{r}
customers <- data.frame(customers, cat_columns)
head(customers)
```

Create data master from categorical and converted categorical columns
```{r}
Gender <- unique(customers[c('Gender','Gender.1')])
Ever_Married <- unique(customers[c('Ever_Married','Ever_Married.1')])
Graduated <- unique(customers[c('Graduated', 'Graduated.1')])
Profession <- unique(customers[c('Profession','Profession.1')])
Spending_Score <- unique(customers[c('Spending_Score','Spending_Score.1')])
Var_1 <- unique(customers[c('Var_1','Var_1.1')])
```

Create a variable for clustering with k-means
```{r}
used_data <- c('Gender.1','Ever_Married.1','Age','Graduated.1','Profession.1','Work_Experience','Spending_Score.1','Family_Size','Var_1.1')
```

Look for optimal cluster for k-means clustering
```{r}
set.seed(100)
sse <- sapply(1:10, 
			  function(param_k)
			  {kmeans(customers[used_data], param_k, nstart=25)$tot.withinss}
			 )
print(sse)

#graphic elbow effect
max_cluster <- 10
ssdata = data.frame(cluster=c(1:max_cluster),sse)
ggplot(ssdata, aes(x=cluster,y=sse)) +
                geom_line(color="red") + geom_point() +
                ylab("Within Cluster Sum of Squares") + xlab("Number of Cluster") +
                geom_text(aes(label=format(round(sse, 2), nsmall = 2)),hjust=-0.2, vjust=-0.5) +
  scale_x_discrete(limits=c(1:max_cluster))

```

Cluster with k-means
```{r}
#k-means function creates 5 clusters with 25 random scenarios and saves them into segmentation
set.seed(100)
segmentation <- kmeans(x = customers[used_data], centers = 5, nstart = 25)
segmentation
```

Merge clustering result into data frame customers
```{r}
customers$Cluster <- segmentation$cluster
str(customers)
```

Check clustering result
```{r}
head(which(customers$Cluster == 1))
length(which(customers$Cluster == 1))
segmentation$ifault
customers[which(customers$Cluster == 1),]
```

Name the segmentation
```{r}
CustomerSegmentation <- data.frame(Cluster = c(1,2,3,4,5), Segmentation = c("Diamond Senior", "Silver Middle", "Silver Young", "Gold Professional", "Diamond Professional"))
```

Combine the references and save to file
```{r}
ClusterID <- list(Gender = Gender, Ever_Married = Ever_Married, Graduated = Graduated, Profession = Profession, Spending_Score = Spending_Score, Var_1 = Var_1, segmentation = segmentation, CustomerSegmentation = CustomerSegmentation, used_data = used_data)
saveRDS(ClusterID, "Cluster ID.rds")
```

Load Cluster Identity from file
```{r}
Cluster.ID <- readRDS(file = 'Cluster ID.rds')
Cluster.ID
```
Test clustering with new data
```{r}
#create new data
new_data <- data.frame(ID = '257157', Gender = 'Male', Ever_Married = 'No', Age = 34, Graduated = 'Yes', Profession = 'Artist', Work_Experience = 6, Spending_Score = 'Low', Family_Size = 1, Var_1 = 'Cat_2')

#merge new data with data from Cluster Identity
new_data <- merge(new_data, Cluster.ID$Gender)
new_data <- merge(new_data, Cluster.ID$Ever_Married)
new_data <- merge(new_data, Cluster.ID$Graduated)
new_data <- merge(new_data, Cluster.ID$Profession)
new_data <- merge(new_data, Cluster.ID$Spending_Score)
new_data <- merge(new_data, Cluster.ID$Var_1)
new_data

#clustering new data
Cluster.ID$CustomerSegmentation[which.min(sapply(1:5, function(x) sum((new_data[Cluster.ID$used_data] - Cluster.ID$segmentation$centers[x,])^2))),]
```

